Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-01-31 src/tex/ms.tex

\begin{abstract}
    The propagation of gravitational waves can reveal fundamental feature of the structure of spacetime.
    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun of parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
    Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
    From 70 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
\end{abstract}

@@

We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{@Thomas cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
We apply a distance prior uniform in comoving volume \mi{@Thomas is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.

@@

\section{Results}
\label{sec:Results}
In this section, we present the results of our study.
We first show the PE results of all events and present the results of Bayesian hierarchical modeling.
Then, we will discuss some special events individually.

\begin{figure*}
    \script{violin_kappa.py}
    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
    \caption{
        The violin plot shows the posterior of $\kappa$ for 69 events in this study.
        Each violin represents a different event.
        The color of the violins represents the quotient of the median and standard deviation of the posterior.
        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
        The pink horizontal solid line represents $\kappa=0$.
    }
    \label{fig:violin_kappa}
\end{figure*}

@@

\subsection{Bayesian hierarchical modeling}
% Corner plot of the mean and standard deviation of kappa
We then show the result of Bayesian hierarchical modeling in figure \ref{fig:corner_Gaussian}.
The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.

>>>>

\begin{abstract}
    The propagation of gravitational waves can reveal the fundamental feature of the structure of spacetime.
    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
    Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
    From 70 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
\end{abstract}

@@

We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, with \textsc{IMRPhenomXPHM} as the reference waveform.
We apply a distance prior uniform in comoving volume \mi{@Thomas is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.

@@

\section{Results}
\label{sec:Results}
In this section, we present the results of our study.
We first show the PE results of all events and present the results of Bayesian hierarchical modeling.
Then, we will discuss some special events individually.
% What special treatment do we give to events

\begin{figure*}
    \script{violin_kappa.py}
    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
    \caption{
        The violin plot shows the posterior of $\kappa$ for 70 events in this study.
        Each violin represents a different event.
        The color of the violins represents the quotient of the median and standard deviation of the posterior.
        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
        The pink horizontal solid line represents $\kappa=0$.
    }
    \label{fig:violin_kappa}
\end{figure*}
% show the excluded events visually different

@@

\subsection{Bayesian hierarchical modeling}
% Corner plot of the mean and standard deviation of kappa
We then show the result of Bayesian hierarchical modeling in figure \ref{fig:corner_Gaussian}.
The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
% talk about what events pull sigma in figure 2
% create a session for all the outliers
